\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)

### 2016-01-23 evaluation.tex
update
updates

\newcommand{\colspacing}{\hspace{1.8em}}
\begin{table}[t]
\small
\centering
\setlength{\tabcolsep}{0.3em}
\begin{tabular}{l rrrrr rr}
\centering
\input{experiments/tables/statistics.tex}
\end{tabular}
\caption{Statistics from modeling the environment of the three Microsoft Azure-based systems under test. The ($\star$) denotes ``awaiting confirmation''.}
\label{tab:stats}
\vspace{-4mm}
\end{table}

@@

Developing both MigratingTable and its \psharp test harness took approximately five person weeks. The harness was developed in parallel with the actual system. This differs from the other two case studies, where the modeling activity occurred independently and after the development process had finished.

@@

We have implemented two different schedulers that are responsible for choosing the next \psharp machine to execute in each scheduling point: a \emph{random} scheduler, which randomly chooses a machine from a list of enabled\footnote{A \psharp machine is considered enabled when it has at least one event in its queue waiting to be dequeued and handled.} machines; and a \emph{randomized priority-based}~\cite{burckhardt2010pct} scheduler, which always schedules the highest priority enabled machine (these priorities change at random points during execution, based on a random distribution). We decided to only use random scheduling, because it has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}.
The random seed for both schedulers was generated using the current time. The priority-based scheduler was further configured with a budget of 2 random machine priority change switches per execution.

@@

For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:migrating} on eleven bugs, including eight \emph{organic} bugs that actually occurred in development and three \emph{notional} bugs (denoted by $*$), which are other code changes that we deemed interesting ways of making the system incorrect. The harness found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}. The remaining four bugs (marked $\diamond$) were not caught with our default test harness in the 100,000 executions (a process that required less than 30 minutes). We believe this is because the inputs and schedules that trigger them are too rare in the used distribution. To confirm this, we wrote a custom test case for each bug with a specific input that triggers it and were able to quickly reproduce the bugs; the table shows the results of these runs.
Note that the random scheduler only managed to trigger seven of the MigratingTable bugs; we had to use the priority-based scheduler to trigger the remaining four bugs.

The QueryStreamedBackUpNewStream bug in MigratingTable, which was found using \psharp, stands out because it reflects a type of oversight that can easily occur as systems evolve. This bug is in the implementation of a \emph{streaming read} from the virtual table, which should return a stream of all rows in the table sorted by key. The essential implementation idea is to perform streaming reads from both backend tables and merge the results. According to the \texttt{IChainTable} specification, each row read from a stream may reflect the state of the table at any time between when the stream was started and the row was read. The developers sketched a proof that the merging process would preserve this property as long as the migrator was only copying rows from the old table to the new table. But when support was added for the migrator to delete the rows from the old table after copying, it became possible for the backend streams to see the deletion of a row from the old table but not its insertion into the new table, even though the insertion happens first, and the row would be missed. The \psharp test discovered this bug in a matter of seconds. The MigratingTable developers spent just 10 minutes analyzing the trace to diagnose what was happening, although admittedly, this was after they added MigratingTable-specific trace information and had several days of experience analyzing traces. Out of the box, \psharp traces include only machine- and event-level information, but it is easy to add application-specific information, and we did so in all of our case studies.

>>>>

\newcommand{\colspacing}{\hspace{1.8em}}
\begin{table}[t]
\small
\centering
\setlength{\tabcolsep}{0.3em}
\begin{tabular}{l rrrrr rr}
\centering
\input{experiments/tables/statistics.tex}
\end{tabular}
\caption{Statistics from modeling the environment of the three Microsoft Azure-based systems under test. The ($\star$) denotes ``awaiting confirmation''.}
\label{tab:stats}
\vspace{-2mm}
\end{table}

@@

Developing both MigratingTable and its \psharp test harness took approximately five person weeks. The harness was developed in parallel with the actual system. This differs from the other two case studies, where the modeling activity occurred independently and after the development process.

@@

We implemented two different schedulers that are responsible for choosing the next \psharp machine to execute in each scheduling point: a \emph{random} scheduler, which randomly chooses a machine from a list of enabled\footnote{A \psharp machine is considered enabled when it has at least one event in its queue waiting to be dequeued and handled.} machines; and a \emph{randomized priority-based}~\cite{burckhardt2010pct} scheduler, which always schedules the highest priority enabled machine (these priorities change at random points during execution, based on a random distribution). We decided to use these two schedulers, because random scheduling has proven to be efficient for finding concurrency bugs~\cite{thomson2014sct, deligiannis2015psharp}. The random seed for the schedulers was generated using the current system time. The priority-based scheduler was configured with a budget of 2 random machine priority change switches per execution.

@@

For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:migrating} on eleven bugs, including eight \emph{organic} bugs that actually occurred in development and three \emph{notional} bugs (denoted by $*$), which are other code changes that we deemed interesting ways of making the system incorrect. The harness found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}. The remaining four bugs (marked $\diamond$) were not caught with our default test harness in the 100,000 executions. We believe this is because the inputs and schedules that trigger them are too rare in the used distribution. To confirm this, we wrote a custom test case for each bug with a specific input that triggers it and were able to quickly reproduce the bugs; the table shows the results of these runs. Note that the random scheduler only managed to trigger seven of the MigratingTable bugs; we had to use the priority-based scheduler to trigger the remaining four bugs.

The QueryStreamedBackUpNewStream bug in MigratingTable, which was found using \psharp, stands out because it reflects a type of oversight that can easily occur as systems evolve. This bug is in the implementation of a \emph{streaming read} from the virtual table, which should return a stream of all rows in the table sorted by key. The essential implementation idea is to perform streaming reads from both backend tables and merge the results. According to the \texttt{IChainTable} specification, each row read from a stream may reflect the state of the table at any time between when the stream was started and the row was read. The developers sketched a proof that the merging process would preserve this property as long as the migrator was only copying rows from the old table to the new table. But when support was added for the migrator to delete the rows from the old table after copying, it became possible for the backend streams to see the deletion of a row from the old table but not its insertion into the new table, even though the insertion happens first, and the row would be missed.

The \psharp test discovered the above bug in a matter of seconds. The MigratingTable developers spent just 10 minutes analyzing the trace to diagnose what was happening, although admittedly, this was after they added MigratingTable-specific trace information and had several days of experience analyzing traces. Out of the box, \psharp traces include only machine- and event-level information, but it is easy to add application-specific information, and we did so in all of our case studies.

### 2016-01-23 intro.tex
updates

To summarize, our contributions are as follows:
\begin{itemize}
\item We present a new methodology for modeling, specifying properties of correctness, and systematically testing real distributed systems with \psharp.
\item We discuss our experience of using \psharp to test three distributed storage systems that are used inside Microsoft, finding subtle bugs that could not be found with traditional testing techniques.
\item We evaluate the cost and benefits of using our approach, and show that \psharp can detect bugs in a small setting and with easy to understand traces.
\end{itemize}

>>>>

To summarize, our contributions are as follows:
\begin{itemize}
\item We present a new methodology for modeling, specifying properties of correctness, and systematically testing real distributed systems with \psharp.
\item We discuss our experience of using \psharp to test three distributed storage systems built on top of Microsoft Azure, finding subtle bugs that could not be found with traditional testing techniques.
\item We evaluate the cost and benefits of using our approach, and show that \psharp can detect bugs in a small setting and with easy to understand traces.
\end{itemize}

### 2016-01-23 overview.tex
updates

In this work, we are proposing a solution between the above two extremes: test the implementation of one or more components against a modeled \psharp environment. The benefit of our approach is that it can detect bugs in the \emph{actual} implementation, but it can do so in a much reduced state-space. Testing with \psharp does not come for free; developers have to invest effort and time into building a test harness using \psharp. However, we argue that developers already spend significant time in building test suites for distributed systems prior to deployment. The \psharp approach augments this effort; by investing time in modeling the environment, it offers dividends by finding more bugs (see \S\ref{sec:eval}). In principle, our methodology is \emph{not specific} to \psharp and the .NET framework, and can be used in combination with any other programming framework that has equivalent capabilities.

@@

\psharp uses object-oriented language features such as interfaces and virtual method dispatch to connect the real code with the modeled code. Programmers in industry are used to working with such features, and heavily employ them in production for testing purposes. In our experience, this significantly lowers the bar for engineering teams inside Microsoft to embrace \psharp for testing.

In \S\ref{sec:overview:safety} and \S\ref{sec:overview:liveness}, we discuss how safety and liveness specifications can be expressed in \psharp to check if the example is correct. The interested reader can refer to the \psharp GitHub repository\footnote{\url{https://github.com/p-org/PSharp}} to find a manual and samples (e.g.\ Paxos~\cite{lamport1998part} and Raft~\cite{ongaro2014raft}), and read \S\ref{sec:vnext}, \S\ref{sec:migrating} and \S\ref{sec:fabric} to learn more details about how \psharp was used to model and test real distributed storage systems in Microsoft.

@@

Liveness property specifications generalize nontermination; a liveness violation is an infinite trace that exhibits lack of progress. Typically, a liveness property is specified via a temporal logic formula~\cite{Pnueli1977, lamport1994temporal}. We take a different approach and allow the programmer to write a \emph{liveness monitor}~\cite{desai2015building}. Similar to a safety monitor, a liveness monitor can receive, but not send, events.

A liveness monitor contains two special types of states: \emph{hot} and \emph{cold}. A hot state denotes a point in the execution where progress is required but has not happened yet; e.g.\ a node has failed but a new one has not come up yet. A liveness monitor enters a hot state when it is notified that the system must make progress. The liveness monitor leaves the hot state and enters a cold state when it is notified that the system has progressed. An infinite execution is erroneous if the liveness monitor is in a hot state for an infinitely long time. Our liveness monitors can encode arbitrary temporal logic properties.
%A thermometer analogy is useful for understanding liveness monitors: a hot state increases the temperature by a small value; a cold state resets the temperature to zero; a liveness error happens if the temperature becomes infinite.
%The distinction between ordinary and cold states is useful for modeling progress for each instance of an infinite stream of events, such as a periodic request or a periodic timer expiration.

A liveness violation is witnessed by an \emph{infinite} execution in which all concurrently executing \psharp machines are \emph{fairly} scheduled. Unfortunately, it is clearly not possible to generate an infinite execution by executing a program for a finite amount of time. Therefore, our implementation of liveness checking in \psharp approximates an infinite execution using several heuristics. In this work, we consider an execution longer than a large user-supplied bound as an ``infinite'' execution~\cite{killian2007life, musuvathi2008fair}. Note that fairness is not relevant when using this heuristic, due to our pragmatic use of a large bound instead of proper handling of infinite executions.

>>>>

In this work, we are proposing a solution between the above two extremes: test the implementation of one or more components against a modeled \psharp environment. The benefit of our approach is that it can detect bugs in the \emph{actual} implementation, and achieve this in a much reduced state-space. Testing with \psharp does not come for free; developers have to invest effort and time into building a test harness using \psharp. However, developers already spend significant time in building test suites for distributed systems prior to deployment. The \psharp approach augments this effort; by investing time in modeling the environment, it offers dividends by finding more bugs (see \S\ref{sec:eval}). In principle, our methodology is \emph{not specific} to \psharp and the .NET framework, and can be used in combination with any other programming framework that has equivalent capabilities.

@@

\psharp uses object-oriented language features such as interfaces and virtual method dispatch to connect the real code with the modeled code. Developers in industry are used to working with such features, and heavily employ them in testing production systems. In our experience, this significantly lowers the bar for engineering teams inside Microsoft to embrace \psharp for testing.

In \S\ref{sec:overview:safety} and \S\ref{sec:overview:liveness}, we discuss how safety and liveness specifications can be expressed in \psharp to check if the example is correct. The details of how \psharp was used to model and test real distributed storage systems in Microsoft are covered in \S\ref{sec:vnext}, \S\ref{sec:migrating} and \S\ref{sec:fabric}. Interested readers can also refer to the \psharp GitHub repository\footnote{\url{https://github.com/p-org/PSharp}} to find a manual and samples (e.g.\ Paxos~\cite{lamport1998part} and Raft~\cite{ongaro2014raft}).

@@

Liveness property specifications generalize nontermination; a liveness violation is an infinite trace that exhibits lack of progress. Typically, a liveness property is specified via a temporal logic formula~\cite{Pnueli1977, lamport1994temporal}. We take a different approach and allow the developers to write a \emph{liveness monitor}~\cite{desai2015building}. Similar to a safety monitor, a liveness monitor can receive, but not send, events.

A liveness monitor contains two special types of states: \emph{hot} and \emph{cold}. A hot state denotes a point in the execution where progress is required but has not happened yet; e.g.\ a node has failed but a new one has not launched yet. A liveness monitor enters a hot state when it is notified that the system must make progress. The liveness monitor leaves the hot state and enters the cold state when it is notified that the system has progressed. An infinite execution is erroneous if the liveness monitor is in the hot state for an infinitely long period of time. Our liveness monitors can encode arbitrary temporal logic properties.
%A thermometer analogy is useful for understanding liveness monitors: a hot state increases the temperature by a small value; a cold state resets the temperature to zero; a liveness error happens if the temperature becomes infinite.
%The distinction between ordinary and cold states is useful for modeling progress for each instance of an infinite stream of events, such as a periodic request or a periodic timer expiration.

A liveness violation is witnessed by an \emph{infinite} execution in which all concurrently executing \psharp machines are \emph{fairly} scheduled. Since it is impossible to generate an infinite execution by executing a program for a finite amount of time, our implementation of liveness checking in \psharp approximates an infinite execution using several heuristics. In this work, we consider an execution longer than a large user-supplied bound as an ``infinite'' execution~\cite{killian2007life, musuvathi2008fair}. Note that checking for fairness is not relevant when using this heuristic, due to our pragmatic use of a large bound.

### 2016-01-23 paper.tex
updates

We thank our shepherd Haryadi Gunawi for his valuable guidance, and the reviewers for their constructive comments. We also thank Ankush Desai from UC Berkeley, and Rich Draves, David Goebel, David Nichols and SW Worth from Microsoft, for their valuable feedback and discussions on various stages of this work. This research was partially supported by a gift from Intel Corporation.

>>>>

We thank our shepherd Haryadi Gunawi for his valuable guidance, and the reviewers for their constructive comments. We also thank Ankush Desai from UC Berkeley, and Rich Draves, David Goebel, David Nichols and SW Worth from Microsoft, for their valuable feedback and discussions at the various stages of this work. We acknowledge that this research was partially supported by a gift from Intel Corporation.

### 2016-01-23 vnext.tex
updates towards camera ready
updates

Microsoft Azure Storage is a cloud storage system that provides customers the ability to store seemingly limitless amounts of data. It has grown from tens of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion~\cite{greenberg2015keynote}.

Azure Storage vNext is the next generation storage system for Microsoft Azure, where the primary design target is to scale the storage capacity by more than 100$\times$. Similar to the current system, vNext employs containers, called \emph{extents}, to store data. Extents can be several gigabytes each, consisting of many data blocks, and are replicated over multiple \emph{Extent Nodes} (ENs). However, in contrast to the current system, which uses a Paxos-based, centralized mapping from extents to ENs, vNext achieves scalability by using a \emph{distributed mapping}. In vNext, extents are divided into partitions, with each partition managed by a light-weight \emph{Extent Manager} (ExtMgr). This partitioning is illustrated in Figure~\ref{fig:vnext}.

@@

One of the responsibilities of an ExtMgr is to ensure that every managed extent maintains enough \emph{replicas} in the system. To achieve this, an ExtMgr receives frequent periodic \emph{heartbeat} messages from every EN that it manages. EN failure is detected by missing heartbeats. An ExtMgr also receives less frequent, but still periodic, \emph{synchronization reports} from every EN that it manages. The sync reports list all the extents (and associated metadata) stored on the EN. Based on these two types of messages, an ExtMgr identifies which ENs have failed, and which extents are affected by the failure and are missing replicas as a result. The ExtMgr then schedules tasks to repair the affected extents and distributes the tasks to the necessary ENs. The ENs then repair the extents from the existing replicas and lazily update the ExtMgr via their next periodic sync reports. All this communication between an ExtMgr and the ENs occurs via network engines installed in each component of vNext (see Figure~\ref{fig:vnext}).

@@

\item \emph{Stress testing}, in which an ExtMgr is launched together with multiple ENs and multiple extents. The test keeps repeating the following process: injecting an EN failure, launching a new EN and checking that the affected extents are eventually repaired.
\end{enumerate}

\noindent
Despite of the extensive testing efforts, the vNext developers were plagued for months by an elusive bug in the ExtMgr logic. All the unit test suites and integration test suites successfully passed on each test run. However, the stress test suite failed \emph{from time to time} after very long executions; in these cases, certain replicas of some extents failed without subsequently being repaired.

@@

  [OnEvent(TimerTickEvent, ProcessExtentRepair)]
  void ProcessExtentRepair() {
    // extent repair loop driven by external timer
    ExtMgr.ProcessEvent(new ExtentRepairEvent());
  }
}
\end{lstlisting}
\vspace{-2mm}
\caption{The real Extent Manager is wrapped inside the \texttt{ExtentManager} \psharp machine.}
\label{fig:wrap_target}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.95\linewidth]{img/extent_manager}
\vspace{-2mm}
\caption{Internal components of the real Extent Manager in Microsoft Azure Storage vNext.}
\label{fig:extentmanager}
\end{figure}

\textbf{Real Extent Manager}.
The real ExtMgr (see Figure~\ref{fig:extentmanager}) contains two data structures: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}. The \texttt{ExtentCenter} maps extents to their hosting ENs. It is updated upon receiving a periodic sync reports from an EN. Recall that a sync report from a particular EN lists all the extents stored at the EN. Its purpose is to update the ExtMgr's possible out-of-date view of the EN with the ground truth. The \texttt{ExtentNodeMap} maps ENs to their latest heartbeat times.

@@

// modeled engine for intercepting Extent Manager messages
class ModelNetEngine : NetworkEngine {
  public override void SendMessage(Socket s, Message msg) {
    // intercept and relay Extent Manager messages
    PSharp.Send(TestingDriver, new ExtMgrMsgEvent(), s, msg);
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Modeled vNext network engine.}
\label{fig:enginecode}
\vspace{-3mm}
\end{figure}

@@

  // extent node failure logic
  [OnEvent(FailureEvent, ProcessFailure)]
  void ProcessFailure() {
    // notifies the monitor that this EN failed
    PSharp.Notify<RepairMonitor>(new ENFailedEvent(), this);
    PSharp.Halt(); // terminate this P# machine
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Code snippet of the modeled EN.}
\label{fig:modeled_en}
\vspace{-2mm}
\end{figure}

@@

  [OnEvent(RepeatedEvent, GenerateTimerTick)]
  void GenerateTimerTick() {
    // nondeterministic choice controlled by P#
    if (PSharp.Nondet())
      PSharp.Send(Target, new TimerTickEvent());
    PSharp.Send(this, new RepeatedEvent()); // loop
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Modeling timer expiration using \psharp.}
\label{fig:modeled_timer}
\vspace{-2mm}
\end{figure}

@@

  void InjectNodeFailure() {
    // nondeterministically choose an EN using P#
    var node = (Machine)PSharp.Nondet(ExtentNodes);    
    PSharp.Send(node, new FailureEvent()); // fail chosen EN
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Code snippet of the TestingDriver machine.}
\label{fig:testing_driver}
\vspace{-2mm}
\end{figure}

@@

  // hot state: repairing
  hot state Repairing {
    [OnEvent(ExtentRepairedEvent, ProcessRepairCompletion)]
    void ProcessRepairCompletion(ExtentNodeMachine en) {
      ExtentNodesWithReplica.Add(en);
      if (ReplicaCount == Harness.REPLICA_COUNT_TARGET)
        jumpto Repaired;
    }
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{The RepairMonitor liveness monitor.}
\label{fig:monitor}
\vspace{-2mm}
\end{figure}

@@

The original \psharp trace did not include sufficient details to allow the developers to identify the root cause of the problem. Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debugging outputs in each iteration. After several iterations, the developers were able to pinpoint the exact culprit and immediately propose a solution for fixing the bug. Once the proposed solution was implemented, the developers ran the test harness again. No bugs were found during 100,000 executions, a process that only took on the order of tens of minutes.

@@

The following is one particular execution sequence resulting in this liveness bug: (i) EN$_0$ fails and is detected by the EN expiration loop; (ii) EN$_0$ is removed from \texttt{ExtentNodeMap}; (iii) \texttt{ExtentCenter} is updated and the replica count drops from 3 (which is the target) to 2; (iv) ExtMgr receives a sync report from EN$_0$; (v) the extent center is updated and the replica count increases again from 2 to 3. This is problematic since the replica count is equal to the target, which means that the extent repair loop will never schedule any repair task. At the same time, there are only two \emph{true} replicas in the system, which is one less than the target. This execution sequence leads to one missing replica; repeating this process two more times would result in all replicas missing, but ExtMgr would still think that all replicas are healthy. In production, such a bug can cause a very serious incident of customer data loss.

>>>>

Microsoft Azure Storage is a cloud storage system that provides customers the ability to store seemingly limitless amounts of data. It has grown from tens of petabytes in 2010 to exabytes in 2015, with the total number of objects stored exceeding 60 trillion~\cite{greenberg2015keynote}.

Azure Storage vNext is the next generation storage system currently being developed for Microsoft Azure, where the primary design target is to scale the storage capacity by more than 100$\times$. Similar to the current system, vNext employs containers, called \emph{extents}, to store data. Extents can be several gigabytes each, consisting of many data blocks, and are replicated over multiple \emph{Extent Nodes} (ENs). However, in contrast to the current system, which uses a Paxos-based, centralized mapping from extents to ENs, vNext achieves scalability by using a \emph{distributed mapping}. In vNext, extents are divided into partitions, with each partition managed by a lightweight \emph{Extent Manager} (ExtMgr). This partitioning is illustrated in Figure~\ref{fig:vnext}.

@@

One of the responsibilities of an ExtMgr is to ensure that every managed extent maintains enough \emph{replicas} in the system. To achieve this, an ExtMgr receives frequent periodic \emph{heartbeat} messages from every EN that it manages. EN failure is detected by missing heartbeats. An ExtMgr also receives less frequent, but still periodic, \emph{synchronization reports} from every EN. The sync reports list all the extents (and associated metadata) stored on the EN. Based on these two types of messages, an ExtMgr identifies which ENs have failed, and which extents are affected by the failure and are missing replicas as a result. The ExtMgr then schedules tasks to repair the affected extents and distributes the tasks to the remaining ENs. The ENs then repair the extents from the existing replicas and lazily update the ExtMgr via their next periodic sync reports. All the communications between an ExtMgr and the ENs occur via network engines installed in each component of vNext (see Figure~\ref{fig:vnext}).

@@

\item \emph{Stress testing}, in which an ExtMgr is launched together with multiple ENs and many extents. The test keeps repeating the following process: injects an EN failure, launches a new EN and checks that the affected extents are eventually repaired.
\end{enumerate}

\noindent
Despite the extensive testing efforts, the vNext developers were plagued for months by an elusive bug in the ExtMgr logic. All the unit test suites and integration test suites successfully passed on each test run. However, the stress test suite failed \emph{from time to time} after very long executions; in these cases, certain replicas of some extents failed without subsequently being repaired.

@@

  [OnEvent(TimerTickEvent, ProcessExtentRepair)]
  void ProcessExtentRepair() {
    // extent repair loop driven by external timer
    ExtMgr.ProcessEvent(new ExtentRepairEvent());
  }
}
\end{lstlisting}
\vspace{-3mm}
\caption{The real Extent Manager is wrapped inside the \texttt{ExtentManager} \psharp machine.}
\label{fig:wrap_target}
\vspace{-1mm}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.95\linewidth]{img/extent_manager}
\vspace{-2mm}
\caption{Internal components of the real Extent Manager in Microsoft Azure Storage vNext.}
\label{fig:extentmanager}
\vspace{-2mm}
\end{figure}

\textbf{Real Extent Manager}.
The real ExtMgr (see Figure~\ref{fig:extentmanager}) contains two data structures: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}. The \texttt{ExtentCenter} maps extents to their hosting ENs. It is updated upon receiving a periodic sync report from an EN. Recall that a sync report from a particular EN lists all the extents stored at the EN. Its purpose is to update the ExtMgr's possible out-of-date view of the EN with the ground truth. The \texttt{ExtentNodeMap} maps ENs to their latest heartbeat times.

@@

// modeled engine for intercepting Extent Manager messages
class ModelNetEngine : NetworkEngine {
  public override void SendMessage(Socket s, Message msg) {
    // intercept and relay Extent Manager messages
    PSharp.Send(TestingDriver, new ExtMgrMsgEvent(), s, msg);
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Modeled vNext network engine.}
\label{fig:enginecode}
\vspace{-2mm}
\end{figure}

@@

  // extent node failure logic
  [OnEvent(FailureEvent, ProcessFailure)]
  void ProcessFailure() {
    // notifies the monitor that this EN failed
    PSharp.Notify<RepairMonitor>(new ENFailedEvent(), this);
    PSharp.Halt(); // terminate this P# machine
  }
}
\end{lstlisting}
\vspace{-5mm}
\caption{Code snippet of the modeled EN.}
\label{fig:modeled_en}
\vspace{-4mm}
\end{figure}

@@

  [OnEvent(RepeatedEvent, GenerateTimerTick)]
  void GenerateTimerTick() {
    // nondeterministic choice controlled by P#
    if (PSharp.Nondet())
      PSharp.Send(Target, new TimerTickEvent());
    PSharp.Send(this, new RepeatedEvent()); // loop
  }
}
\end{lstlisting}
\vspace{-5mm}
\caption{Modeling timer expiration using \psharp.}
\label{fig:modeled_timer}
\vspace{-2mm}
\end{figure}

@@

  void InjectNodeFailure() {
    // nondeterministically choose an EN using P#
    var node = (Machine)PSharp.Nondet(ExtentNodes);    
    PSharp.Send(node, new FailureEvent()); // fail chosen EN
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Code snippet of the TestingDriver machine.}
\label{fig:testing_driver}
\vspace{-4mm}
\end{figure}

@@

  // hot state: repairing
  hot state Repairing {
    [OnEvent(ExtentRepairedEvent, ProcessRepairCompletion)]
    void ProcessRepairCompletion(ExtentNodeMachine en) {
      ExtentNodesWithReplica.Add(en);
      if (ReplicaCount == Harness.REPLICA_COUNT_TARGET)
        jumpto Repaired;
    }
  }
}
\end{lstlisting}
\vspace{-5mm}
\caption{The RepairMonitor liveness monitor.}
\label{fig:monitor}
\vspace{-2mm}
\end{figure}

@@

The original \psharp trace did not include sufficient details to allow the developers to identify the root cause of the problem. Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debugging outputs in each iteration. After several iterations, the developers were able to pinpoint the exact culprit and immediately propose a solution for fixing the bug. Once the proposed solution was implemented, the developers ran the test harness again. No bugs were found during 100,000 executions, a process that only took a few minutes.

@@

The following is one particular execution sequence resulting in this liveness bug: (i) EN$_0$ fails and is detected by the EN expiration loop; (ii) EN$_0$ is removed from \texttt{ExtentNodeMap}; (iii) \texttt{ExtentCenter} is updated and the replica count drops from 3 (which is the target) to 2; (iv) ExtMgr receives a sync report from EN$_0$; (v) the extent center is updated and the replica count increases again from 2 to 3. This is problematic since the replica count is equal to the target, which means that the extent repair loop will never schedule any repair task. At the same time, there are only two \emph{true} replicas in the system, which is one less than the target. This execution sequence leads to one missing replica; repeating this process two more times would result in all replicas missing, but ExtMgr would still think that all replicas are healthy. If released to production, this bug would have caused a very serious incident of customer data unavailability.
